---
title: "Analysis 2_Decision trees_stats"
output: rmarkdown::github_document
---

This code calculates the chi square statistics for each split in the decision trees by quartile for the N and More tasks.  

It utilize these files in the Data folder:

* data_n_tree_stats_helper.csv
* data_more_tree_stats_helper.csv
Both files contain columns: quartile, split_node, left_terminal, right_terminal, feature, feature_value.

The value are hand-coded after producing the tree plots using Analysis 2_Decision trees_plots.Rmd, that contains inner node split information

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load libraries & data
```{r, message=FALSE}
library(party)
library(tidyverse)
library(here)
library(caret)
#library(partykit) 
```
## Load data (only need to change this for the More task analysis)
```{r}
# indicate the task name, make it easier for adopting the same code for both the N and M tasks
task_name = "more"

# input files
data_long = read.csv(here(paste0("Data/data_", task_name,"_long.csv")))

data_tree_stats_help = read.csv(here(paste0("Data/data_", task_name, "_tree_stats_helper.csv")))
# this file contain information about each split information (e.g., left/right terminal nodes, the splitting features, left/right feature_value)

data_item_binomial = read.csv(here(paste0("Data/data_", task_name, "_item_binomial.csv")))

# within each quartile, some items don't have at least 10 children responses. exclude them from the item level accuracy graph, but not the overall analysis???
min_item_n = 10

data_item_binomial = data_item_binomial %>% 
  filter(item_total_n >= min_item_n) 
```

## Grow Ctrees
```{r}
set.seed(240)

# make sure that all feature cols are factors
data_long = data_long %>%
  mutate(bigger_places = as.factor(bigger_places),
         places = as.factor(places),
         length_diff = as.factor(length_diff),
         one_digit_diff_not_zero = as.factor(one_digit_diff_not_zero),
         transposition = as.factor(transposition),
         inserting_zero = as.factor(inserting_zero))

Analysis2_chisq_results = NULL

set.seed(240)

for (i in 1:4) {
  # get quartile data
  temp_data = subset(data_long, quartile == i) 
  
  temp_splits = subset(data_tree_stats_help, quartile == i)
  temp_split_nodes = unique(temp_splits$split_node)
  temp_split_nodes_num = length(temp_split_nodes)
  
  # grow ctree
  ct = ctree(factor(acc) ~ 
               bigger_places + places + length_diff + one_digit_diff_not_zero +
               transposition + inserting_zero, data = temp_data)
  
  plot(ct)
  
  # save tree terminal node number to quartile data
  a = where(ct)
  temp_data_with_node_number = cbind(temp_data,a)
  
  # loop through each split, get all left and right split data, and computer the stats
  for (j in 1:temp_split_nodes_num) {
    temp_data = subset(temp_splits, split_node == temp_split_nodes[j])
    
    left_nodes = na.omit(unique(subset(temp_splits, split_node == temp_split_nodes[j])$left_terminal))
    left_data = temp_data_with_node_number %>%
      filter(a %in% left_nodes) %>%
      mutate(group = "left")
    
    right_nodes = na.omit(unique(subset(temp_splits, split_node == temp_split_nodes[j])$right_terminal))
    right_data = temp_data_with_node_number %>%
      filter(a %in% right_nodes) %>%
      mutate(group = "right")
    
   
    left_acc = mean(left_data$acc)
    right_acc = mean(right_data$acc)
    
    all_data = rbind(left_data, right_data)
    chisq_results = chisq.test(all_data$group, all_data$acc)
    
    # retrieve features
    feature = unique(subset(temp_splits, split_node == temp_split_nodes[j])$feature)
    
    left_feature_value = subset(temp_splits, split_node == temp_split_nodes[j]) %>%
      filter(!is.na(left_terminal)) %>%
      select(feature_value)
    
    right_feature_value = subset(temp_splits, split_node == temp_split_nodes[j]) %>%
      filter(!is.na(right_terminal)) %>%
      select(feature_value)
    
    # save the results
    Analysis2_chisq_results_temp = data.frame("quartile" = i, "split" = temp_split_nodes[j], "left_acc" = left_acc, "right_acc" = right_acc,"chisq_value" = chisq_results$statistic, "chisq_df" = chisq_results$parameter, "chisq_p" = chisq_results$p.value, "feature" = feature, "left_feature_value" = unique(left_feature_value$feature_value), "right_feature_value" = unique(right_feature_value$feature_value))
    
    Analysis2_chisq_results = rbind(Analysis2_chisq_results, Analysis2_chisq_results_temp)
  }
}

write.csv(Analysis2_chisq_results, here("Data/Analysis2_chisq_results.csv"), row.names = FALSE)
```
